---
title: "Assignment 9"
output:
  html_document:
    toc: true
    toc_float: true
    collapsed: false
    number_sections: false
    toc_depth: 1
    #code_folding: hide
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(message=FALSE,warning=FALSE, cache=TRUE)
```

#Use simulations to determine the effect of the following on the t-value (from an independent sample t-test): 

##The sample size
As sample size increases, the absolute value of the t statistic increases. The graph seems to show the opposite because the t-value is increasing negatively. 
```{r}
sample_1_std <- 20
sample_2_std <- 20
sample_1_mean <- 70
sample_2_mean <- 75
ts_vs_size <- c() 
n_exps <- 100
sample_size_vector <- c(2:200)
for (ind in 1:length(sample_size_vector)){
  sample_sizes <- sample_size_vector[ind] 
  t_values <- c()
  for (t_ind in 1:n_exps){
    sample_1_scores <- rnorm(n=sample_sizes, mean=sample_1_mean, sd=sample_1_std)
    sample_2_scores <- rnorm(n=sample_sizes, mean=sample_2_mean, sd=sample_1_std)
    t_values[t_ind] <- t.test(sample_1_scores, sample_2_scores, var.equal = TRUE)$statistic
    }
  ts_vs_size[ind] <- mean(t_values)
}


library(ggplot2)

my_data <- data.frame(value=ts_vs_size, size=sample_size_vector)

ggplot(my_data, aes(x=size, y=value)) +
  geom_point()+ 
  geom_line()

```


##The standard deviations of the estimates
As the standard deviation increases the absolute value of the t statistic decreases. The graph seemingly shows the opposite because the t-value is decreasing negatively (getting smaller).
```{r}

sample_size <- 20
sample_1_mean <- 70
sample_2_mean <- 75
ts_vs_std <- c() 
n_exps <- 100
sample_std_vector <- c(2:20)
for (ind in 1:length(sample_std_vector)){
  sample_std <- sample_std_vector[ind] 
  ts <- c()
  for (t_ind in 1:n_exps){
    sample_1_scores <- rnorm(n=sample_size, mean=sample_1_mean, sd=sample_std)
    sample_2_scores <- rnorm(n=sample_size, mean=sample_2_mean, sd=sample_std)
    ts[t_ind] <- t.test(sample_1_scores, sample_2_scores, var.equal = TRUE)$statistic
    }
  ts_vs_std[ind] <- mean(ts) 
}


library(ggplot2)

my_data <- data.frame(value=ts_vs_std, std=sample_std_vector)

ggplot(my_data, aes(x=std, y=value)) +
  geom_point()+ 
  geom_line()

```


##The effect size
As the effect size increases, the t-value increases. 
```{r}

sample_size <- 20
sample_1_mean <- 70
sample_2_mean <- 70
sample_1_std <- 20
sample_2_std <- 20
ts_vs_ef <- c() 
n_exps <- 100
sample_es_vector <- c(2:100)
for (ind in 1:length(sample_es_vector)){
  sample_effect <- sample_es_vector[ind] 
  ts <- c()
  for (t_ind in 1:n_exps){
    sample_1_scores <- rnorm(n=sample_size, mean=sample_1_mean+sample_effect, sd=sample_1_std)
    sample_2_scores <- rnorm(n=sample_size, mean=sample_2_mean, sd=sample_2_std)
    ts[t_ind] <- t.test(sample_1_scores, sample_2_scores, var.equal = TRUE)$statistic
  }
  ts_vs_ef[ind] <- mean(ts) 
}


library(ggplot2)

my_data <- data.frame(value=ts_vs_ef, effect=sample_es_vector)

ggplot(my_data, aes(x=effect, y=value)) +
  geom_point()+ 
  geom_line()

```

#Use the same sample size and standard deviation for both groups. Which of those seem to increase the t-value?
A larger sample size increases the absolute value of the t-value because you are increasing the amplification of the effect size. A larger standard deviation decreases the absolute value of the t-value because the error is increased. 
```{r}
#Large sample size, small sd
sample_size <- 100
sample_1_mean <- 70
sample_2_mean <- 75
sample_1_std <- 10
sample_2_std <- 10
sample_1_scores <- rnorm(n=sample_size, mean=sample_1_mean, sd=sample_1_std)
sample_2_scores <- rnorm(n=sample_size, mean=sample_2_mean, sd=sample_2_std)
t.test(sample_1_scores, sample_2_scores, var.equal = TRUE)$statistic


#Large sample size, big sd
sample_size <- 100
sample_1_mean <- 70
sample_2_mean <- 75
sample_1_std <- 30
sample_2_std <- 30
sample_1_scores <- rnorm(n=sample_size, mean=sample_1_mean, sd=sample_1_std)
sample_2_scores <- rnorm(n=sample_size, mean=sample_2_mean, sd=sample_2_std)
t.test(sample_1_scores, sample_2_scores, var.equal = TRUE)$statistic


#Small sample size, small sd
sample_size <- 20
sample_1_mean <- 70
sample_2_mean <- 75
sample_1_std <- 10
sample_2_std <- 10
sample_1_scores <- rnorm(n=sample_size, mean=sample_1_mean, sd=sample_1_std)
sample_2_scores <- rnorm(n=sample_size, mean=sample_2_mean, sd=sample_2_std)
t.test(sample_1_scores, sample_2_scores, var.equal = TRUE)$statistic


#Small sample size, big sd
sample_size <- 20
sample_1_mean <- 70
sample_2_mean <- 75
sample_1_std <- 30
sample_2_std <- 30
sample_1_scores <- rnorm(n=sample_size, mean=sample_1_mean, sd=sample_1_std)
sample_2_scores <- rnorm(n=sample_size, mean=sample_2_mean, sd=sample_2_std)
t.test(sample_1_scores, sample_2_scores, var.equal = TRUE)$statistic

```

