---
title: "Assignment 2"
output:
  html_document:
    toc: true
    toc_float: true
    collapsed: false
    number_sections: false
    toc_depth: 1
    #code_folding: hide
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(message=FALSE,warning=FALSE, cache=TRUE)
```

#Find all flights that:
```{r}
library(dplyr)
library(ggplot2)
install.packages('nycflights13')
library(nycflights13)
```

##Had an arrival delay of two or more hours
```{r}

twohr_delay <- flights %>%
  select(flight, dep_delay) %>%
  filter(dep_delay >= 120)

```


##Flew to Dallas (DAL or DFW)
```{r}
Dallasflights <- flights %>%
  select(flight, dest) %>%
  filter(dest == "DFW")
```

##Were operated by United (UA), American (AA), or Delta (DL) 
```{r}
carriertype <- flights %>%
  select(carrier, flight) %>%
  filter(carrier == "UA" | carrier == "AA" | carrier =="DL")
```

##Departed in summer (July, August, and September)
```{r}
summerflights <- flights %>%
  select(month, flight) %>%
  filter(month==7:9)
```

##Arrived more than two hours late, but didn’t leave late
```{r}
latefights <- flights %>%
  select(flight, dep_delay, arr_delay) %>%
  filter(dep_delay==0 & arr_delay > 120)
```

##Were delayed by at least an hour, but arrival delay is less than 30 minutes.
```{r}
delayedflights <- flights %>%
  select(flight, dep_delay, arr_delay) %>%
  filter(dep_delay >=60 & arr_delay < 30)

```

##Departed before 6am (remember to check the time)
```{r}
earlyflights <- flights %>%
  select(flight, dep_time) %>%
  filter(dep_time < 600)
```


#Is there a relationship between proportion of cancelled flights and the average delays? Remember that a missing dep_delay or arr_delay (i.e., is.na() is TRUE) represents a cancelled flight. Make a scatter plot that have the proportion of cancelled flights on the x-axis, and the average delay on the y-axis. Group the observations by month. Also make sure you plot straight line using geom_smooth(method='lm').
```{r}


canceled <- flights %>% 
  group_by(month) %>%
  summarize(count_canceled=sum(is.na(dep_delay)), 
            av_delay = mean(dep_delay, na.rm = TRUE))
  
ggplot(canceled, aes(x=count_canceled, y=av_delay))+
  geom_point(aes(color=month))+
  geom_smooth(method='lm')


```

