---
title: "Assignment 10"
output:
  html_document:
    toc: true
    toc_float: true
    collapsed: false
    number_sections: false
    toc_depth: 1
    #code_folding: hide
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(message=FALSE,warning=FALSE, cache=TRUE)
```



#1) Conduct a small power analysis to determine the minimum N needed to acheive a power of 80% using a correlation test (i.e.,cor.test(x,y)). To do this, you need to use random numbers from rnorm.

```{r}
N <- seq(3,1000,2)
n_exp <- 100
n_vs_power <- c()

for(this_n in 1:length(N)) {
  sample_size <- N[this_n]
  sig_pval <- c()
  
  for (exp in 1:n_exp){
  var10 <- rnorm(sample_size)
  var11 <- var10 + rnorm(sample_size, sd=10)
  corr_test <- cor.test(var10,var11)
  sig_pval[exp] <- corr_test$p.value < 0.05
  }
  n_vs_power[this_n] <- mean(sig_pval)
}

library(ggplot2)
ggplot(data.frame(N,n_vs_power), aes(x=N, y=n_vs_power))+
  geom_line()+
  geom_smooth()


min(N[n_vs_power == min(n_vs_power[n_vs_power>0.8])])


```

#2) Rˆ2 is called the coefficient of determination and it is usually used as a measure of the effect size. Conduct another power analysis that shows the numimum effect size needed to achieve 80% power using N = 20.
```{r}

N <- 20
n_exp <- 100
rsq_vs_power <- c()
rsq <- c(0, 1, 0.01)

for(this_n in 1:length(rsq)) {
  this_rsqu <- rsq[this_n]
  sig_rsq <- c()
  
  for (exp in 1:n_exp){
  var10 <- rnorm(N)
  var11 <- var10 + rnorm(N, sd=10)
  corr_test <- cor.test(var10,var11)
  sig_rsq[exp] <- corr_test$estimate
  }
  rsq_vs_power[this_n] <- mean(sig_rsq)
}

library(ggplot2)
ggplot(data.frame(N,rsq_vs_power), aes(x=N, y=rsq_vs_power))+
  geom_line()+
  geom_smooth()


min(rsq[rsq_vs_power == min(rsq_vs_power[rsq_vs_power>0.8])])



power_vs_effect_size


```



```{r}
power_vs_effect_size <- c()
effect_size <- seq(0,1,0.01)

for (effect_size_ind in 1:length(effect_size)) {
  this_effect_size <- effect_size[effect_size_ind]
  sample_size <- 20
  pop_mean <- 70
  pop_sd <- 20
  alpha <- 0.05
  n_runs <- 100
  significance <- c()
  for (run in 1:n_runs) {
    grp_experimental <- rnorm(n=sample_size, mean=pop_mean+this_effect_size, sd=pop_sd)
    grp_control <- rnorm(n=sample_size, mean=pop_mean, sd=pop_sd)
    rsq <- cor.test(grp_experimental, grp_control)$estimate
    significance[run] <- rsq
  }
  
}

library(ggplot2)
my_df <- data.frame(effect_size=effect_size, power=power_vs_effect_size)

ggplot(my_df, aes(x=effect_size, y=power))+
  geom_line()+
  geom_point()


```
