---
title: "Exam 1"
output:
  html_document:
    toc: true
    toc_float: true
    collapsed: false
    number_sections: false
    toc_depth: 1
    #code_folding: hide
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(message=FALSE,warning=FALSE, cache=TRUE)
```

#1) Write a command that tells us many days where 5 customers have waited (one line).

```{r, eval=FALSE}
waiting_customers <- c(4,2,1,1,0,2,1,2,4,2,5,3,1,5,1,5,1,2,1,1,3,4,2,4,3)


table(waiting_customers)[6]


```

#2) Write a command that tells us the maximum of those numbers
```{r, eval=FALSE}
max(waiting_customers)
```


#3) Write a command that tells us the mean of those numbers
```{r, eval=FALSE}
mean(waiting_customers)
```

#4)  Write a command that prints the standard deviation of those numbers.
```{r, eval=FALSE}
sd(waiting_customers)
```

#5) Using the following matrix:
```{r, eval=FALSE}
numbers <- seq(from=1, to=100,by=2)

numbers_matrix <- matrix(numbers, nrow=5, ncol=10)
```

Write the command that tells us the dimensions of this matrix (number of rows, number of columns).

```{r, eval=FALSE}

dim(numbers_matrix)

```

#6) Using the following matrix:

```{r, eval=FALSE}
numbers <- seq(from=1, to=100,by=2)
numbers_matrix <- matrix(numbers, nrow=5, ncol=10)
```

#7) Write the command that calculates the mean of each column, without using the `mean` function (i.e., use a for-loop)

```{r, eval=FALSE}
mean_function <-function(x_mean){ 
  sum(x_mean)/length(x_mean)} 

  print(mean_function(x_mean))
  
  
column_1 <- numbers_matrix[,1]
mean_function(column_1)
column_2 <- numbers_matrix[,2]
mean_function(column_2)
column_3 <- numbers_matrix[,3]
mean_function(column_3)
column_4 <- numbers_matrix[,4]
mean_function(column_4)
column_5 <- numbers_matrix[,5]
mean_function(column_5)
column_6 <- numbers_matrix[,6]
mean_function(column_6)
column_7 <- numbers_matrix[,7]
mean_function(column_7)
column_8 <- numbers_matrix[,8]
mean_function(column_8)
column_9 <- numbers_matrix[,9]
mean_function(column_9)
column_10 <- numbers_matrix[,10]
mean_function(column_10)


```


#8) Write the command that calculates the sum of each column, without using the `sum` function (i.e., use a for-loop)

```{r, eval=FALSE}
numbers <- seq(from=1, to=100,by=2)
numbers_matrix <- matrix(numbers, nrow=5, ncol=10)

j <- 0
for (numbers in numbers_matrix){
j <- numbers + numbers
}
print(j)


```


#9) Write the command that calculates the sum of each column, without using the `sum` function (i.e., use a for-loop)

```{r, eval=FALSE}

numbers <- seq(from=1, to=100,by=2)
numbers_matrix <- matrix(numbers, nrow=5, ncol=10)

total_sum <- 0
for (x in numbers_matrix) {
  total_sum <- numbers_matrix[x,] + numbers_matrix[x+c(1:4),] 
}
print(total_sum)
```

```{r, eval=FALSE}
Linda_over_years <- babynames[babynames$name == "Linda" & babynames$sex == "F",]
Max_Linda_over_years <- babynames[babynames$name == "Linda" & babynames$sex == "F" & babynames$n == max(Linda_over_years$n),]
Max_Linda_over_years$year
```



