---
title: "Exam 2"
output:
  html_document:
    toc: true
    toc_float: true
    collapsed: false
    number_sections: false
    toc_depth: 1
    #code_folding: hide
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(message=FALSE,warning=FALSE, cache=TRUE)
```


```{r, eval=FALSE}
library(dplyr)
library(ggplot2)
install.packages('nycflights13')
library(nycflights13)
```

#1) Make a histogram of the air_time in ggplot. Try using different values for bins to see if that will change your histogram.

```{r, eval=FALSE}

ggplot(data=flights, aes(x=air_time))+
  geom_histogram(bins=10)


ggplot(data=flights, aes(x=air_time))+
  geom_histogram(bins=50)


ggplot(data=flights, aes(x=air_time))+
  geom_histogram(bins=100)


```


#2) Calculate how many canceled flights per day? Make a bar plot for canceled flights per day.

```{r, eval=FALSE}

canceled <- flights %>% 
  group_by(day) %>%
  summarize(count_canceled=sum(is.na(dep_delay)))


ggplot(canceled, aes(x=day, y=count_canceled))+
  geom_bar(stat = "identity")

```

#3) Which destinations have more canceled flights? Make a list of the worst 10 destinations and the best 10 destinations in terms of the number of canceled flights.

```{r, eval=FALSE}

worst_dest <- flights %>%
  group_by(dest) %>%
  summarize(count_canceled=sum(is.na(dep_delay))) %>%
  arrange(desc(count_canceled)) %>%
  mutate(count=1:length(dest)) %>%
  filter(count <= 10)

best_dest <- flights %>%
  group_by(dest) %>%
  summarize(count_canceled=sum(is.na(dep_delay))) %>%
  arrange(count_canceled) %>%
  mutate(count=1:length(dest)) %>%
  filter(count <= 10)

```


#4) Which destinations have higher average delayed flights in terms of the dep_delay? Choose the worst 10 and make them into a bar graph.
```{r, eval=FALSE}

high_delay <- flights %>%
  group_by(dest) %>%
  filter(!is.na(dep_delay)) %>%
  summarize(av_delay=mean(dep_delay)) %>%
  arrange(desc(av_delay)) %>%
  mutate(count=1:length(dest)) %>%
  filter(count <= 10)

ggplot(high_delay, aes(x=dest, y=av_delay)) +
  geom_bar(stat = "identity")

```

#5) Which flights are the fastest flights? (to get that, you need to calculate speed. Speed is distance divided by air_time*60 so you get a value in mph)

```{r, eval=FALSE}

flightdata <- flights


fast_flights <- flights %>%
  mutate(speed=distance/air_time*60) %>%
  arrange(desc(speed)) %>%
  mutate(count=1:length(dest)) %>%
  filter(count <= 10) %>%
  select(dest, speed)
 
```


#6) Which carrier has the worst delays? Since we have 16 unique carriers, make a barplot of the 16 carriers along with their average delays. Use a different color for each bar.

```{r, eval=FALSE}

carrier_delays <- flights %>%
  group_by(carrier) %>%
  filter(!is.na(dep_delay)) %>%
  summarize(av_delay=mean(dep_delay)) %>%
  arrange(desc(av_delay)) %>%
  mutate(Longest_Delays_Ranked=1:length(carrier))

ggplot(carrier_delays, aes(x=carrier, y=av_delay))+
  geom_bar(stat = "identity", aes(fill=carrier))


```


#7) What time of day should you fly if you want to avoid delays as much as possible? You should make a barplot (or a line point) with the time of the day on the x-axis and the average delays on the y-axis.
```{r, eval=FALSE}

daydelay <- flights %>%
  select(day,dep_time,dep_delay) %>%
  group_by(dep_time) %>%
  summarize(mean_delay=mean(dep_delay, na.rm=TRUE)) %>%
  mutate(time=dep_time/100, delay_time= mean_delay/100) 
  

ggplot(daydelay, aes(x=dep_time, y=mean_delay))+
  geom_line()+
  ggtitle("Average Delays in One Day") +
      xlab("Time of Day (in military time)") +
      ylab("Average Delay") 

```


#8) If you want to go to Chicago (code: `ORD`), what NYC airport and which carrier you should consider so that you have the minimum delay possible?
```{r, eval=FALSE}

chicagoflights <- flights %>%
  select(origin, dest, dep_delay) %>%
  filter(dest=="ORD") %>%
  group_by(origin) %>%
  summarize(av_delay=mean(dep_delay, na.rm=TRUE))


```

#9) What are the furthest destinations and how long does it take to get there? (make a list of 10 furthest destinations along with the flight duration in hours)
```{r, eval=FALSE}

furthestdest <- flights %>%
  select(distance, air_time, dest) %>%
  group_by(dest) %>%
  summarize(av_distance=mean(distance)) %>%
  arrange(desc(av_distance)) %>%
  mutate(Top_10=1:length(av_distance)) %>%
  filter(Top_10 <= 10)
  

```

#10) What are the most common destinations from NYC airports (i.e., destinations that have highest number of flights)? Make a list of 10.
```{r, eval=FALSE}
most_common_dest <- flights %>%
  select(dest) %>%
  
  
  
ggplot(most_common_dest, aes(x=dest)) +
  geom_histogram(bin=20)

```

#11) Which NYC airport and which carrier has more long flights? Can you guess from this list which carriers have lots of international flights? (group_by both NYC airports and carrier).
```{r, eval=FALSE}

```

