---
title: "Final Project"
output:
  html_document:
    toc: true
    toc_float: true
    collapsed: false
    number_sections: false
    toc_depth: 1
    #code_folding: hide
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(message=FALSE,warning=FALSE, cache=TRUE)
```

#Loading the data
```{r}

school_data<- read.csv('/Users/AnaFranz/Downloads/2018-2019_School_Demographic_Snapshot.csv')

?read.csv



library(tidyverse)
school_data <- read_csv('~/Downloads/2018-2019_School_Demographic_Snapshot.csv')

  
  
  
```


#What was the school with the highest total enrollement in 2018-19 year?
```{r}
library(dplyr)
# convert year to a number by keeping only first 4 charachters in the year string 
school_data$Year <- substr(school_data$Year, 1, 4)
school_data$Year <- as.numeric(school_data$Year)

# read a single column
school_data$`Total Enrollment`

  average_enrollement_2018 <- school_data %>%
    filter(Year=='2018') %>%
    select(`School Name`, Year, `Total Enrollment`) %>%
    arrange(`Total Enrollment`) %>%
    top_n(10)
  
  average_enrollement_2014 <- school_data %>%
    filter(Year=='2014') %>%
    select(`School Name`, Year, `Total Enrollment`) %>%
    arrange(`Total Enrollment`) %>%
    top_n(10)
  
  average_enrollement_2015 <- school_data %>%
    filter(Year=='2015') %>%
    select(`School Name`, Year, `Total Enrollment`) %>%
    arrange(`Total Enrollment`) %>%
    top_n(10)
  
  
  
```


#Which schools do not have a high school?
```{r}

 nohighschool <- school_data %>%
    filter(`Grade 9`==0 & `Grade 10`==0 & `Grade 11` ==0 & `Grade 12`==0) %>%
    select(`School Name`) %>%
    distinct()
    
```

#Correlation between % in poverty and % of minitories (black and hispanic) present at school
```{r}
school_data$`% Hispanic` <- substr(school_data$`% Hispanic`, 1, 4)
school_data$`% Hispanic` <- as.numeric(school_data$`% Hispanic`)

minoritieshisp <- school_data$`% Hispanic`

school_data$`% Black` <- substr(school_data$`% Black`, 1, 4)
school_data$`% Black` <- as.numeric(school_data$`% Black`)

minoritiesblk <- school_data$`% Black`

school_data$`% Poverty` <- substr(school_data$`% Poverty`, 1, 4)
school_data$`% Poverty` <- as.numeric(school_data$`% Poverty`)

poverty <- school_data$`% Poverty`

cor(poverty, minoritieshisp)

cor(poverty, minoritiesblk)

library(dplyr)

minorities <- school_data %>%
  filter(Year=='2018') %>%
  select(`School Name`, `Economic Need Index`, `% Black`, `% Asian`, `% Hispanic`) %>%
  mutate(`% Minorities`= `% Black` + `% Asian`+`% Hispanic`) %>%
  arrange(`% Minorities`) %>%
  select(`School Name`, `% Minorities`,`Economic Need Index`)

percentminorities <- minorities$`% Minorities`
economicneedindex <- minorities$`Economic Need Index`

cor( percentminorities, economicneedindex)

library(ggplot2)
ggplot(minorities, aes(x=`% Minorities`, y=`Economic Need Index`))+
  geom_point()+
  geom_smooth(method='lm') +
  ggtitle("Percent Minorities vs Economic Need Index",) +
  xlab("Percent of Minorities in School") +
  ylab("Economic Need Index of School") 

  
         
minoritiesschool <- school_data %>%
  filter(Year=='2018') %>%
  select(`School Name`, `% Black`, `% Asian`, `% Hispanic`) %>%
  mutate(`% Minorities`= `% Black` + `% Asian`+`% Hispanic`) %>%
  arrange(desc(`% Minorities`)) %>%
  select(`School Name`, `% Minorities`) %>%
  filter(`% Minorities`> 0.500)

economicneedindexschool <- school_data %>%
  filter(Year=='2018') %>%
  select(`School Name`, `% Black`, `% Asian`, `% Hispanic`, `Economic Need Index`) %>%
  mutate(`% Minorities`= `% Black` + `% Asian`+`% Hispanic`) %>%
  arrange(desc(`% Minorities`)) %>%
  select(`School Name`, `% Minorities`, `Economic Need Index`) %>%
  filter(`Economic Need Index`> 0.500)

minoritiesasian <- school_data %>%
  filter(Year=='2018') %>%
  select(`School Name`, `% Asian`) %>%
  arrange(desc(`% Asian`)) %>%
  filter(`% Asian`> 0.500)

brooklyn <- c(1,1,1,1,1,1)
brooklyn <-sum(brooklyn)
manhattan <- c(1,1,1,1,1)
MAN <- SUM(manhattan)
bronx <- c(1,1,1,1,1,1,1,1,1,1,1,1,1,1,1,1,1,1,1,1)
BX <- sum(bronx)
queens <- c(1,1)
QN <- sum(queens)

boroughdata <- data.frame(Borough=c("Brooklyn","Manhattan","Bronx","Queens"),
                          Number_of_Schools= c(6,4,20,2))

?data.frame

barplot(as.numeric(boroughdata))

ggplot(boroughdata, aes(x=Borough, y=Number_of_Schools))+
  geom_bar(stat = 'identity', aes(fill=Borough)) +
  ggtitle("Location of Schools with Minorities as Majority",) +
  xlab("Percent of Minorities in School") +
  ylab("Number of Schools") 

library(ggplot2)

ggplot(my_df, aes(x=BK, MAN, BX, QN))+
         geom_bar(stat = 'identity')

barplot(as.numeric(boroughdata), names.arg = c("a","b","c","d"))

```


#Which schools have the highest percent of female students in 2018? Which has the lowest? (not including all girls or all boys schools).
```{r}

  highestfemale <- school_data %>%
    filter(Year=='2018') %>%
    select(`School Name`, `% Female`) %>%
    filter(!`% Female`== 1 & !`% Female`== 0 ) %>%
    arrange(desc(`% Female`)) %>%
    top_n(15) 
 


library(ggplot2)
ggplot(highestfemale, aes(x=`School Name`, y=`% Female`)) +
  geom_bar(stat = 'identity')+
  theme(axis.text.x = element_text(face="bold", color="#993333", 
                           size=7, angle=45),
          axis.text.y = element_text(face="bold", color="#993333", 
                           size=7, angle=45))
  
  lowestfemale <- school_data %>%
    filter(Year=='2018') %>%
    select(`School Name`, `% Female`) %>%
    filter(!`% Female`== 1 & !`% Female`== 0 ) %>%
    arrange(`% Female`) %>%
    top_n(-15)
  
  
  
```

#Which schools are all boys schools? Which are all girls schools? 
```{r}

  allgirls <- school_data %>%
    select(`School Name`, `% Female`) %>%
    filter(`% Female`== 1) %>%
    distinct()
  
  allboys <- school_data %>%
    select(`School Name`, `% Male`) %>%
    filter(`% Male`== 1) %>%
    distinct()
```




```{r}
library(tidyverse)
library(dplyr)
final_data <- read_csv('~/Downloads/2018-2019_School_Demographic_Snapshot.csv')
final_data$Year <- substr(final_data$Year, 1, 4)
final_data$Year <- as.numeric(final_data$Year)

# read a single column
final_data$Year
final_data$`Economic Need Index`
final_data$`# Poverty`
final_data$`School Name`


#What are the dropout rates between middle-school and highschool?

final_data %>%
  group_by(`School Name`) %>%
  summarize(junior_high_school = mean(c(`Grade 6`,`Grade 7`, `Grade 8`), , na.rm = TRUE),
            high_school = mean(c( `Grade 9`,`Grade 10`,`Grade 11`, `Grade 12`), na.rm = TRUE)) %>%
  filter(junior_high_school!=0 & high_school!=0) %>%
  mutate(dropout =  high_school-junior_high_school )

#Is there a correlation between poverty and dropout rates

#Creating a new data set
#Table Avg and Drop Rates
Dropout_rates_Avgs<-(final_data %>%
                       group_by(`School Name`) %>%
                       summarize(junior_high_school = mean(c(`Grade 6`,`Grade 7`, `Grade 8`), na.rm = TRUE),
                                 high_school = mean(c( `Grade 9`,`Grade 10`,`Grade 11`, `Grade 12`), na.rm = TRUE)) %>%
                       filter(junior_high_school!=0 & high_school!=0) %>%
                       mutate(dropout =  junior_high_school-high_school ))

#Joining Poverty and Dropout rates data set
poverty_df <- final_data %>%
  filter(`School Name` %in% Dropout_rates_Avgs$`School Name`) %>%
  select(`# Poverty`, `School Name`) %>%
  full_join(Dropout_rates_Avgs, by=c('School Name')) %>%
  group_by(`School Name`) %>%
  summarise( drop_mean = mean(dropout),
             poverty_mean = mean(`# Poverty`)) %>%
  filter(drop_mean >0)

# Scatterplot/ Graph to see the correlation
library(ggplot2)
ggplot(poverty_df, aes(x=drop_mean , y= poverty_mean)) +  
  geom_point() +
  geom_smooth(method ='lm')+
  ggtitle("Correlation Between Poverty and Drop-out Rates") +
  xlab("Drop-Out Rate") +
  ylab("Number of Students in Poverty") 


#Checking to see if there a Pearson's correlation betwwen poverty and dropout?
cor(poverty_df$drop_mean, poverty_df$poverty_mean)
cor.test(poverty_df$drop_mean, poverty_df$poverty_mean)

# Correlation between total enrollment and dropout
T_En <- final_data %>%
  group_by(`School Name`) %>%
  summarize(junior_high_school = mean(c(`Grade 6`,`Grade 7`, `Grade 8`), na.rm = TRUE),
            high_school = mean(c( `Grade 9`,`Grade 10`,`Grade 11`, `Grade 12`), na.rm = TRUE),
            avg_e = mean(`Total Enrollment`)) %>%
  filter(junior_high_school!=0 & high_school!=0) %>%
  mutate(dropout = junior_high_school-high_school)

library(ggplot2)
ggplot(T_En, aes(x=dropout , y= avg_e)) +  
  geom_point() +
  geom_smooth(method='lm')+
  ggtitle("Correlation Between Drop-out Rate and Total Enrollment") +
  xlab("Drop-Out Rate") +
  ylab("Total Enrollment") 
```


```{r}
minoritiesschool2014<- school_data %>%
  filter(Year=='2014') %>%
  select(`School Name`, `% Black`, `% Asian`, `% Hispanic`) %>%
  mutate(`% Minorities`= `% Black` + `% Asian`+`% Hispanic`) %>%
  arrange(desc(`% Minorities`)) %>%
  select(`School Name`, `% Minorities`) %>%
  filter(`% Minorities`> 0.500)

minoritiesschool2015<- school_data %>%
  filter(Year=='2015') %>%
  select(`School Name`, `% Black`, `% Asian`, `% Hispanic`) %>%
  mutate(`% Minorities`= `% Black` + `% Asian`+`% Hispanic`) %>%
  arrange(desc(`% Minorities`)) %>%
  select(`School Name`, `% Minorities`) %>%
  filter(`% Minorities`> 0.500)

minoritiesschool2016<- school_data %>%
  filter(Year=='2016') %>%
  select(`School Name`, `% Black`, `% Asian`, `% Hispanic`) %>%
  mutate(`% Minorities`= `% Black` + `% Asian`+`% Hispanic`) %>%
  arrange(desc(`% Minorities`)) %>%
  select(`School Name`, `% Minorities`) %>%
  filter(`% Minorities`> 0.500)


minoritiesschool2017<- school_data %>%
  filter(Year=='2017') %>%
  select(`School Name`, `% Black`, `% Asian`, `% Hispanic`) %>%
  mutate(`% Minorities`= `% Black` + `% Asian`+`% Hispanic`) %>%
  arrange(desc(`% Minorities`)) %>%
  select(`School Name`, `% Minorities`) %>%
  filter(`% Minorities`> 0.500)

minoritiesschool2018<- school_data %>%
  filter(Year=='2018') %>%
  select(`School Name`, `% Black`, `% Asian`, `% Hispanic`) %>%
  mutate(`% Minorities`= `% Black` + `% Asian`+`% Hispanic`) %>%
  arrange(desc(`% Minorities`)) %>%
  select(`School Name`, `% Minorities`) %>%
  filter(`% Minorities`> 0.500)



economicneedindexschool <- school_data %>%
  filter(Year=='2018') %>%
  select(`School Name`, `% Black`, `% Asian`, `% Hispanic`, `Economic Need Index`) %>%
  mutate(`% Minorities`= `% Black` + `% Asian`+`% Hispanic`) %>%
  arrange(desc(`% Minorities`)) %>%
  select(`School Name`, `% Minorities`, `Economic Need Index`) %>%
  filter(`Economic Need Index`> 0.500)

```




```{r}

samplingsd <- data.frame(Probability=c(1/3, 2/9, 2/9, 2/9),
                         Value=c(0, 1.4142,3.5355, 4.9497))

library(ggplot2)
ggplot(samplingsd, aes(x=Value, y=Probability))+
  geom_bar(stat = 'identity')


```

